Playing Atari Breakout with Keras and R

### 2018-07-15 article/article.Rmd
adding the rest of the code

The most important thing to note here is that `action_model` and `score_model` share all weights. This is important because we will not directly train the `score model`. Instead we will train the `action_model` using gradient descent and use the `score_model` just to get predictions of rewards for each action.

>>>>

The most important thing to note here is that `action_model` and `score_model` share all weights. This is important because we will not directly train the `score_model`. Instead we will train the `action_model` using gradient descent and use the `score_model` just to get predictions of rewards for each action.

Now we will define a function that will perform a step of the gradient descent algorithm to update the models weights.

```{r}
train_step <- function(models, experience, n_actions) {

  # sampling a minibatch of size 32 from the experience buffer
  c(s_t, s_t1, terminal, action, reward) %<-% experience$sample(32)

  # calculating the expected reward for s_t1
  expected_reward <- predict(models$score_model, s_t1) %>% apply(1, max)

  # calculating the Q value for state_t
  y <- reward + (!terminal)*gamma*expected_reward

  # transforming the action to a one-hot vector
  action <- to_categorical(action, n_actions)

  # perform the weights update with gradient descent
  loss <- train_on_batch(
    models$action_model, 
    x = list(s_t, action),
    y = y
  )

  # returns the obtained loss invisibly
  invisible(loss)
}
```

## Final loop

Now that we have all auxiliary functions we are ready to implement the final loop. First let's define all the hyperparameters that will control the training phase.

```{r}
observe <- 500 # number of episodes to observe before starting training
explore <- 4000 # nuber of episodes to do exploration
episodes <- 5000 # total number of episodes
replay_memory <- 100000 # number of transitions to remember
environment_name <- "BreakoutDeterministic-v4" # name of the gym environment 
initial_epsilon <- 1 # initial probability of choosing a random action
final_epsilon <- 0.1 # after `explore` iterations what should be the probability of choosing a rondom action
gamma <- 0.99 # gamma from q-learning
input_shape <- c(53, 40, 4) # size of the pre-processed array
```

Now let's start our objects:

```{r}
env <- gym$make(environment_name)
experience <- Experience$new(replay_memory)
history <- rep(NA, episodes)
models <- build_model(input_shape, env$action_space$n)
epsilon <- 1L
```

And run the final loop:

```{r}
for (i in seq_len(episodes)) {

  if (i >= observe & i < (observe + explore)) 
    epsilon <- epsilon - (initial_epsilon/final_epsilon) / explore

  if (i >= observe)
    train <- TRUE
  else 
    train <- FALSE

  time <- system.time({
    history[i] <- play_episode(env, epsilon, models, experience, train)  
  })

  last100 <- tail(history[!is.na(history)], 100) %>% mean() %>% round(2)
  time <- round(time[3], 2)

  # save model weights
  if(i %% 500 == 0)
    save_model_weights_hdf5(
      models$score_model, 
      "model_weights.hdf5", 
      overwrite = TRUE
    )

  cat(glue::glue("Episode: {i} | Score: {history[i]} | Mean: {last100} | Elapsed: {time} sec"), "\n")
}
```

Note, this loop can take a lot of time. It takes ~24h in a MacBook Air.
